---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# magicr

<!-- badges: start -->
<!-- badges: end -->

The goal of `magicr` is to enable analysis of Magic the Gathering data in R.

This package is in **early development**: functionality is likely to change!

## Installation

You can install the development version of `magicr` like so:

``` r
remotes::install_github("joelnitta/magicr")
```

## Example

Download some draft game data for *Khans of Tarkir* from [17lands.com](https://www.17lands.com)

```{r example}
library(magicr)
mr_get_17lands_data(
  "KTK", "game", "premier",
  # only load a small portion of the data for this demo
  nrows = 10, select = 1:8)
```

## Copyright Notice

`magicr` may include unofficial Fan Content permitted under the [Wizards of the
Coast Fan Content Policy](https://company.wizards.com/en/legal/fancontentpolicy).
Any literal and graphical information presented via this package and its
documentation pages about _Magic: The Gathering_, including card images, the
mana symbols, and Oracle text, is copyright of Wizards of the Coast, LLC, a
subsidiary of Hasbro, Inc. This software is not produced by, endorsed by,
supported by, or affiliated with Wizards of the Coast, LLC or
[17lands](https://www.17lands.com).

## License

[MIT](LICENSE)