---
title: "Downloading files and using the cache"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Downloading files and using the cache}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

One of the main purposes of `magicr` is to obtain data about drafts from [17lands.com](https://17lands.com).
These typically come in the form of rather large, compressed CSV files.
`magicr` helps you maintain these files by storing them in a folder referred to as the "cache".

Let's see how to use this feature.

## Location of the cache

```{r setup}
library(magicr)
```

```{r move-cache, include = FALSE}
library(fs)

# Temporarily move any files in the cache to a temp dir
cache_files <- mr_cache_dir() |>
  dir_ls()
temp_dir <- path(tempdir(), "magicr_cache")
if (dir_exists(temp_dir)) dir_delete(temp_dir)
dir_create(temp_dir)
file_move(cache_files, temp_dir)
```

The location of the cache defaults to a location specific to your OS that is designated for storing files related to this package.
You can see where this is located like this:

```{r cache-loc}
mr_cache_dir()
```

Notice that the exact path will differ based on your computer setup.

You can check the current contents of the cache like so:

```{r cache-loc-ls}
list.files(mr_cache_dir())
```

## Downloading files to the cache

By default, datasets from [17lands](https://17lands.com) will be downloaded to the cache.

Here, let's download *Khans of Tarkir* game data for Premier Draft to the cache.

```{r cache-download}
mr_download_17lands_file(set = "KTK", data_type = "game", event_type = "premier")
list.files(mr_cache_dir())
```

If you prefer to store the files in a folder of your choosing, you can do that with the `cache_dir` argument:

```{r cache-download-custom}
mr_download_17lands_file(
  set = "KTK", data_type = "game", event_type = "premier",
  cache_dir = tempdir())
```

## Cached files won't be re-downloaded

The whole point of having a cache is to avoid re-downloading files.
So if we run `mr_download_17lands_file()` for the same dataset again, it will just return the path to the existing file instead of re-downloading it (you should notice the command runs almost instantaneously this time, whereas the first time it took several seconds to download).

```{r cache-download-again}
mr_download_17lands_file(set = "KTK", data_type = "game", event_type = "premier")
```

However, sometimes new data may be available and you **do** want to re-download.
In that case, set `overwrite = TRUE` to re-download and overwite the existing file:

```{r cache-download-overwrite}
mr_download_17lands_file(
  set = "KTK", data_type = "game", event_type = "premier",
  overwrite = TRUE)
```

## Files are downloaded automatically when loading into R

The `mr_download_17lands_file()` function is user-facing, but if your goal is to load the data directly into R (which is most likely the case!), you can skip this function and use `mr_get_17lands_data()` instead.
`mr_get_17lands_data()` will download the necessary data file for you first, then load it.
And of course it will use the cached file if one is available.

```{r cache-download-load}
ktk <- mr_get_17lands_data(set = "KTK", data_type = "game", event_type = "premier")

# The dataset is large, so just print out part of it
dim(ktk)
ktk[1:6, 1:6]
```

## Coming soon

I plan to add more functionality to see what is in the cache and clear it out or update the files etc.

```{r put-cache-back, include = FALSE}
# Put the cached files back
cache_files_in_temp_dir <- dir_ls(temp_dir)
file_move(cache_files_in_temp_dir, mr_cache_dir())
```
